---
title: "Final"
output: html_document
---

```{r}
load("/Users/mostafaragheb/Downloads/brfss_data/BRFSS2013_a.RData")
```

```{r}
data3.1 <- na.omit(dat3)
```

```{r}
d_Smoke <- data.frame(model.matrix(~ data3.1$X_SMOKER3))
d_Gender <- data.frame(model.matrix(~data3.1$SEX))
d_employ <- data.frame(model.matrix(~data3.1$EMPLOY1))
d_activitylevel <- data.frame(model.matrix(~data3.1$X_PACAT1))
```

```{r}
Data_Final = data.frame(data3.1$BMI_measure , data3.1$X_AGE80 , d_Gender[,2],
                        data3.1$SLEPTIM1, data3.1$d_overweight , d_Smoke[,2:4] ,
                        d_employ[,2:8],d_activitylevel)
names(Data_Final) = c("BMI" , "Age" ,"Gender", "Sleeptime"  , "Overweight",
                      "Smoker" , "use.to.smoke" ,"Never.smoked", "Self.employed",
                      "Out.of.work.for.a.year.or.more",
                      "Out.of.work.less.than.a.year","Homemaker","student",
                      "Retired","Unable.to.work","Active","insufficent.active","Inactive")
```

```{r}
Model_1 = lm(BMI ~ Age + Gender + Sleeptime + Smoker + use.to.smoke + Never.smoked 
             + Self.employed + Out.of.work.for.a.year.or.more +
               Out.of.work.less.than.a.year + Homemaker + student + Retired +
               Unable.to.work  + insufficent.active + Inactive, data = Data_Final)
plot(Model_1$fitted.values , Data_Final$BMI)
    abline(a=coef(Model_1)[1],
           b=coef(Model_1)[2],
           col="red",
           lwd=3)
    
summary(Model_1)

cor(Data_Final$BMI , Data_Final$Age)
cor(Data_Final$BMI , Data_Final$Sleeptime)

```

```{r}
attach(Data_Final)
NNobs <- length(BMI)
set.seed(12345)
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(Data_Final,graph_obs)  

plot(BMI ~ Sleeptime, pch = 16, col = rgb(0.5, 0.5, 0.5), data = dat_graph)

to_be_predicted1 <- data.frame(Data_Final) 
to_be_predicted1$yhat <- predict(Model_1, newdata = to_be_predicted1)

lines(yhat ~ Sleeptime, data = to_be_predicted1)

```

```{r}
Model_2 = lm((BMI^2) ~ Age + Gender + Sleeptime + Smoker + use.to.smoke + Never.smoked 
             + Self.employed + Out.of.work.for.a.year.or.more +
               Out.of.work.less.than.a.year + Homemaker + student + Retired +
               Unable.to.work + insufficent.active + Inactive, data = Data_Final)
plot(Data_Final$BMI,Model_2$fitted.values)
    abline(a=coef(Model_2)[1],
           b=coef(Model_2)[2],
           col="red",
           lwd=3)
    
summary(Model_2)
```
```{r}
Model_3 = lm( log(BMI) ~  Age + Gender + Sleeptime + Smoker + use.to.smoke +
                Never.smoked  + Self.employed + Out.of.work.for.a.year.or.more +
               Out.of.work.less.than.a.year + Homemaker + student + Retired +
               Unable.to.work + insufficent.active + Inactive , data = Data_Final)

    plot(Data_Final$BMI,Model_3$fitted.values)
    abline(a=coef(Model_3)[1],
           b=coef(Model_3)[2],
           col="red",
           lwd=3)

summary(Model_3)
```

```{r}
Model_4 = lm(log(Data_Final$BMI) ~ Data_Final$Sleeptime)
summary(Model_4)

plot(Data_Final$Sleeptime,Data_Final$BMI)

```

```{r}
Model_5 = lm(log(Data_Final$BMI) ~ Data_Final$Age)
summary(Model_5)

plot(Data_Final$Age,Data_Final$BMI)
```
```{r}
Model_6 = lm(Data_Final$BMI ~ Data_Final$Sleeptime + Data_Final$Age)
summary(Model_6)

par(mfrow=c(1,2))
plot(Data_Final$Sleeptime,Data_Final$BMI)
abline( a = coef(Model_6)[1] , b = coef(Model_6)[2] , col = "red")
plot(Data_Final$Age,Data_Final$BMI)
abline( a = coef(Model_6)[1] , b = coef(Model_6)[2] , col = "red")
```

```{r}
Model_7 = lm(log(Data_Final$BMI) ~ Data_Final$Sleeptime + Data_Final$Age)
summary(Model_7)

par(mfrow=c(1,2))
plot(Data_Final$Sleeptime,Data_Final$BMI)
abline( a = coef(Model_7)[1] , b = coef(Model_7)[2] , col = "red")
plot(Data_Final$Age,Data_Final$BMI)
abline( a = coef(Model_7)[1] , b = coef(Model_7)[2] , col = "red")
```

```{r}
Model_8 = lm(log(BMI) ~  Sleeptime + use.to.smoke
              + Out.of.work.for.a.year.or.more +
               Homemaker  + insufficent.active, data = Data_Final)
plot(Model_8$fitted.values , Data_Final$BMI)
    abline(a=coef(Model_8)[1],
           b=coef(Model_8)[2],
           col="red",
           lwd=3)
summary(Model_8)    
```

```{r}
require("standardize")
set.seed(654321)
NN <- length(Data_Final$Overweight)
restrict_1 <- (runif(NN) < 0.1) 
summary(restrict_1)
dat_train <- subset(Data_Final, restrict_1)
dat_test <- subset(Data_Final, !restrict_1)
sobj <- standardize(Overweight ~ Age 
                    + Gender + Sleeptime + Smoker + use.to.smoke + Never.smoked 
             + Self.employed + Out.of.work.for.a.year.or.more +
               Out.of.work.less.than.a.year + Homemaker + student + Retired +
               Unable.to.work  + insufficent.active + Inactive + BMI, dat_train, family = binomial)

s_dat_test <- predict(sobj, dat_test)
summary(s_dat_test)
```

```{r}
model_lpm1 <- lm(sobj$formula, data = sobj$data)
summary(model_lpm1)
pred_vals_lpm <- predict(model_lpm1, s_dat_test)
pred_model_lpm1 <- (pred_vals_lpm > 0.5)
table(pred = pred_model_lpm1, true = dat_test$Overweight)
```
#Model lpm1 is showing that the model predicted People Not over weight are 580 wih error of 126, givig the percentage error equal: 17.8% While people over weight are 1519 with an error of 161, giving the percentage error equal 9.5%
```{r}
126/(580+126)
161/(1519+161)
```


```{r}
model_logit1 <- glm(sobj$formula, family = binomial, data = sobj$data)
summary(model_logit1)
pred_vals <- predict(model_logit1, s_dat_test, type = "response")
pred_model_logit1 <- (pred_vals > 0.5)
table(pred = pred_model_logit1, true = dat_test$Overweight)
```

# Model logit1 is showing that the model predicted People NOT Overweight are 705 wih error of 86, givig the percentage error equal: 10.8% While people over weight are 1559 with an error of 36, giving the percentage error equal 2.25%
```{r}
86/(705+86)
36/(36+1559)
```


```{r}
require('randomForest')
set.seed(54321)
model_randFor <- randomForest(as.factor(Overweight) ~ ., data = sobj$data, importance=TRUE, proximity=TRUE)
print(model_randFor)
round(importance(model_randFor),2)
varImpPlot(model_randFor)
pred_model1 <- predict(model_randFor,  s_dat_test)
table(pred = pred_model1, true = dat_test$Overweight)
```
# According to forest model the error percent is:

#Predicted people who are not overweight: 0.01190476
#Predicted people who are overweight:0.01570681
