---
title: "Homework 5, Lab 5, Econ B2000"
author: 'Mostafa Ragheb, Group Members: Akimawe Kadiri, Nicole Kerrison, Christopher Tinevra, Charles Reed'
date: "11/3/2020"
output: github_document
---

```{r}
load("~/Library/Mobile Documents/com~apple~CloudDocs/Economics Master/Econometrics/Labs/acs2017_ny.RData")
suppressMessages(attach(acs2017_ny))
use_varb <- (AGE >= 25) & (AGE <= 65) & (LABFORCE == 2) & (WKSWORK2 > 4) & (UHRSWORK >= 35) & (female == 1) 
dat_use <- subset(acs2017_ny,use_varb)
detach()
```

Model 1 illustrates the linear regression of income wages based on education level,  age, and ethnicity.

```{r}
suppressMessages(attach(dat_use))
model1 <- lm(INCWAGE ~ AGE + educ_hs + educ_college + educ_advdeg + white + AfAm + Hispanic)
summary(model1)
suppressMessages(require(stargazer))
stargazer(model1, type = "text")
detach()

NNobs <- length(INCWAGE)
set.seed(12345)
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

plot(INCWAGE ~ AGE, pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)

to_be_predicted1 <- data.frame(AGE = 25:65, female = 1,white = 1, Hispanic = 0, AfAm =0, educ_hs = 0, educ_college = 1, educ_advdeg = 0)
to_be_predicted1$yhat <- predict(model1, newdata = to_be_predicted1)

lines(yhat ~ AGE, data = to_be_predicted1)
```

The coefficient estimates for Model 1 linear regression demonstrates multiple income wages depending on educaltion level (high school degree, college degree, Advance degree) & ethnicity (white, African American & Hispanics). The samplefor this regression includes females with ages ranging from 25 to 65 years old and are currently part of the labor force and are fully employed with (35 hour work week). The graph aboveshows a positive correlation for income wages as the age increases, and regression line refers to the predicted values taking into consideration white females with a college degree.   


Model 2 illustrates the linear regression of income wages based on education level, age, age squared, and ethnicity.

```{r}
suppressMessages(attach(dat_use))
model2 <- lm(INCWAGE ~ AGE + I(AGE^2) + educ_hs + educ_college + educ_advdeg + white + AfAm + Hispanic)
summary(model2)
suppressMessages(require(stargazer))
stargazer(model2, type = "text")

NNobs <- length(INCWAGE)
set.seed(12345)
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

plot(INCWAGE ~ AGE, pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)

to_be_predicted2 <- data.frame(AGE = 25:65, female = 1, white = 1, Hispanic = 0, AfAm = 0,educ_hs = 0, educ_college = 1, educ_advdeg = 0)
to_be_predicted2$yhat <- predict(model2, newdata = to_be_predicted2)

lines(yhat ~ AGE, data = to_be_predicted2)

detach()
```

The coefficient estimates for Model 2 linear regression continues to demonstrate multiple income wages depending on educaltion level (high school degree, college degree, Advance degree), ethnicity (white, African American & Hispanics) but also taking into consideration age squared. The sample for this regression includes females with ages ranging from 25 to 65 years old and are currently part of the labor force and are fully employed with (35 hour work week). The graph above shows a positive correlation for income wages as the age increases, and regression line refers to the predicted values taking into consideration white females with a college degree. However the regression slope of the graph is positive as age increases, then decreases, indicating diminishing returns for wages as workers reach retirement age (62 years old).

Model 3 illustrates the linear regression of income wages based on education level, log of age, log of age squared, and ethnicity.

```{r}
suppressMessages(attach(dat_use))
model3 <- lm(INCWAGE ~ log(AGE) + I(log(AGE^2)) + educ_hs + educ_college + educ_advdeg + white + AfAm + Hispanic)
summary(model3)
suppressMessages(require(stargazer))
stargazer(model3, type = "text")
detach()

NNobs <- length(INCWAGE)
set.seed(12345)
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

plot(INCWAGE ~ (AGE), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)

to_be_predicted3 <- data.frame(AGE = 25:65, female = 1, white = 1, Hispanic =0, AfAm = 0,educ_hs = 0, educ_college = 1, educ_advdeg = 0)
to_be_predicted3$yhat <- predict(model3, newdata = to_be_predicted3)

lines(yhat ~ AGE, data = to_be_predicted3)
```

Model 3 expands on the conditions of model 2 but this time, the regression model is based on log of age and log of age squared. However, there is no reason to use Log on Age or Log on Age^2 given that age is a discrete variable mostly on intervals between 1 to 100. Log of a given number (x) is the exponent to which another fixed number base (b), must be raised, to produce that number (x). Therefore, it is not practical to present age in log terms. Log expression is more sutable to simplyfy large numbers like income wages. 


Model 4 illustrates the linear regression of income wages based on education level, age, polynomial of ages (Age^exp 2,3,4), and ethnicity.

```{r}
suppressMessages(attach(dat_use))
model4 <- lm(INCWAGE ~ AGE + I(AGE^2) + I(AGE^3) + I(AGE^4) +I(AGE^5) + I(AGE^6) + educ_hs + educ_college+ educ_advdeg + white + AfAm + Hispanic)

summary(model4)
suppressMessages(require(stargazer))
stargazer(model4, type = "text")
detach()

NNobs <- length(INCWAGE)
set.seed(12345)
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

plot(INCWAGE ~ (AGE), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,150000), data = dat_graph)
to_be_predicted4 <- data.frame(AGE = 25:65, female = 1, white = 1, Hispanic =0, AfAm = 0,educ_hs = 0, educ_college = 1, educ_advdeg = 0)
to_be_predicted4$yhat <- predict(model4, newdata = to_be_predicted4)

lines(yhat ~ AGE, data = to_be_predicted4)
```

Model 4 expands on the conditions of model 2 but this time, the regression model includes age as a polynomial terms where age variable raised by different exponents (2,3,4). Displaying age as a polynomial shows during the predicted analysis that as age increases, wages will increase and flatten out at retirement. If we express the dummy variables in terms of polynomials it will not be effective in the regression because dummy variables are binary and only provide values of zero and one.


Lastly, Model 5 expands on the conditions of model 2 but this time, the regression model is based on log of dependent variable which is income wages. 

```{r}
suppressMessages(attach(dat_use))
model5 <- lm(log1p(INCWAGE) ~ AGE + I(AGE^2) + I(AGE^3) + I(AGE^4) + educ_hs + educ_college + educ_advdeg + white + AfAm + Hispanic)
summary(model5)
suppressMessages(require(stargazer))
stargazer(model5, type = "text")
detach()

NNobs <- length(log1p(INCWAGE))
set.seed(12345)
graph_obs <- (runif(NNobs) < 0.1)
dat_graph <-subset(dat_use,graph_obs)  

plot(log1p(INCWAGE) ~ (AGE), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(10,11), data = dat_graph)

to_be_predicted5 <- data.frame(AGE = 25:65, female = 1, white = 1, Hispanic=0, AfAm = 0,educ_hs = 0, educ_college = 1, educ_advdeg = 0)
to_be_predicted5$yhat <- predict(model5, newdata = to_be_predicted5)

lines(yhat ~ AGE, data = to_be_predicted5)
```

Model 5 expands on the conditions of model 2 but this time, the regression model is based on log of income wages which is the dependent variable. Since, Log expression is more sutable to simplyfy large numbers like income wages, the log values will be expressed within a smaller range compared to the original values. In our case, there would not be high values for the regression line given that the Log of income ranges between 10 to 11. Similar to the results from Model 2 regression line, the regression slope for Model 5 shows a positive slope as age increases, the income wages also increase but then decreases indicating a diminishing returns for wages as workers reach retirement age (62 years old).

The table below demonstrates the estimates for the coefficient variables for all the models that were run above (Models 1-5). The estimates coefficient values shows an increase of income wages for female workers (either white, hispanics or African American) depending on education level ranging from High school degree to college degree to Advance degree and also taking into consideration age. The reversing the variables is not ideal due to age and education level do not depend on different income wages. 

``` {r}
suppressMessages(require(stargazer))
stargazer(model1, model2, model3, model4, model5, type = "text")
```
