---
title: "HW 7"
output: html_document
Author: Mostafa Ragheb Team: Kadiri
---

```{r}
load("/Users/mostafaragheb/Desktop/CCNY/4th ( Fall 2020)/Economatrics/Data/NHIS_2014.RData")
```


```{r}
data_use1$earn_lastyr <- as.factor(data_use1$ERNYR_P)
levels(data_use1$earn_lastyr) <- c("0","$01-$4999","$5000-$9999","$10000-$14999","$15000-$19999","$20000-$24999","$25000-$34999","$35000-$44999","$45000-$54999","$55000-$64999","$65000-$74999","$75000 and over",NA,NA,NA)
```

#Filtering data by age over 18 years old
```{r}
attach(data_use1)
use_varb <-  (AGE_P>18)  
dat_use <- subset(data_use1,use_varb)
detach()
```

#Changing Region, Region born and personal health to a dummy variables that can be used in analysis.
```{r}
d_region <- data.frame(model.matrix(~ dat_use$REGION))
d_region_born <- data.frame(model.matrix(~ factor(dat_use$region_born)))
d_PH <- data.frame((model.matrix(~factor(dat_use$person_healthstatus))))
```

#The Model is examining the people over 18 years old, education level , race , gender, martial status, region and place of birth and their health status. 
```{r}
model_logit1 <- glm(NOTCOV ~ AGE_P + I(AGE_P^2) + female + AfAm + Asian + RaceOther  
                    + Hispanic + educ_hs + educ_smcoll + educ_as + educ_bach + educ_adv 
                    + married + widowed + divorc_sep + veteran_stat + REGION + person_healthstatus +  region_born,
                    family = binomial, data = dat_use)

```

#Creating a new data set to include the new dummy variables (region, region born and personal health status)
```{r}
dat_for_analysis_sub <- data.frame(
  dat_use$NOTCOV,
  dat_use$AGE_P,
  dat_use$female,
  dat_use$AfAm,
  dat_use$Asian,
  dat_use$RaceOther,
  dat_use$Hispanic,
  dat_use$educ_hs,
  dat_use$educ_smcoll,
  dat_use$educ_as,
  dat_use$educ_bach,
  dat_use$educ_adv,
  dat_use$married,
  dat_use$widowed,
  dat_use$person_healthstatus,
  dat_use$divorc_sep,
  d_region[,2:4],
  d_region_born[,2:12],
  d_PH[,2:7])

names(dat_for_analysis_sub) <- c("NOTCOV",
                                 "Age",
                                 "female",
                                 "AfAm",
                                 "Asian",
                                 "RaceOther",
                                 "Hispanic",
                                 "educ_hs",
                                 "educ_smcoll",
                                 "educ_as",
                                 "educ_bach",
                                 "educ_adv",
                                 "married",
                                 "widowed",
                                 "Vgood",
                                 "Good",
                                 "Fair",
                                 "Poor",
                                 "Refused",
                                 "Not.ascertained",
                                 "divorc_sep",
                                 "Region.Midwest",
                                 "Region.South",
                                 "Region.West",
                                 "born.Mex.CentAm.Carib",
                                 "born.S.Am",
                                 "born.Eur",
                                 "born.f.USSR",
                                 "born.Africa",
                                 "born.MidE",
                                 "born.India.subc",
                                 "born.Asia",
                                 "born.SE.Asia",
                                 "born.elsewhere",
                                 "born.unknown")


```

```{r}
install.packages("standaridize")
```


#standardizing data set
```{r}
require("standardize")
set.seed(654321)
NN <- length(dat_for_analysis_sub$NOTCOV)
restrict_1 <- (runif(NN) < 0.1) 
summary(restrict_1)
dat_train <- subset(dat_for_analysis_sub, restrict_1)
dat_test <- subset(dat_for_analysis_sub, !restrict_1)
sobj <- standardize(NOTCOV ~ Age + female + AfAm + Asian + RaceOther + Hispanic + 
                      educ_hs + educ_smcoll + educ_as + educ_bach + educ_adv + 
                      married + widowed + Vgood + Good + Fair + Poor + Refused +
                      Not.ascertained + divorc_sep + 
                      Region.Midwest + Region.South + Region.West + 
                      born.Mex.CentAm.Carib + born.S.Am + born.Eur + born.f.USSR + 
                      born.Africa + born.MidE + born.India.subc + born.Asia + 
                      born.SE.Asia + born.elsewhere + born.unknown, dat_train, family = binomial)

s_dat_test <- predict(sobj, dat_test)
summary(s_dat_test)
```


```{r}
model_lpm1 <- lm(sobj$formula, data = sobj$data)
summary(model_lpm1)
pred_vals_lpm <- predict(model_lpm1, s_dat_test)
pred_model_lpm1 <- (pred_vals_lpm > 0.5)
table(pred = pred_model_lpm1, true = dat_test$NOTCOV)
```
# Model lpm1 is showing that the model predicted People with health coverage are 63,195 wih error of 10,006, givig the percentage error equal: 13.6%
```{r}
10006/(63195+10006)
```
#While the predicted value for people who doesn't have coverage is 751 with error of 293, giving the percentage error equal: 28%
```{r}
293/(751+293)
```



```{r}
model_logit1 <- glm(sobj$formula, family = binomial, data = sobj$data)
summary(model_logit1)
pred_vals <- predict(model_logit1, s_dat_test, type = "response")
pred_model_logit1 <- (pred_vals > 0.5)
table(pred = pred_model_logit1, true = dat_test$NOTCOV)
```
# Model logit1 is showing that the model predicted People with health coverage are 62,222 wih error of 8,873, givig the percentage error equal: 12.4%
```{r}
8873/(62222+8873)
```
#While the predicted value for people who doesn't have coverage is 1,884 with error of 1,266, giving the percentage error equal: 40%
```{r}
1266/(1266+1884)
```



```{r}
require('randomForest')
set.seed(54321)
model_randFor <- randomForest(as.factor(NOTCOV) ~ ., data = sobj$data, importance=TRUE, proximity=TRUE)
print(model_randFor)
round(importance(model_randFor),2)
varImpPlot(model_randFor)
# look at confusion matrix for this too
pred_model1 <- predict(model_randFor,  s_dat_test)
table(pred = pred_model1, true = dat_test$NOTCOV)
```
# According to forest model the error percent is:

#Predicted people who has coverage error: 0.01127927
#Predicted people who doesn't have coverage error:0.88953975

```{r}
require(e1071)
# tuned_parameters <- tune.svm(as.factor(NOTCOV) ~ ., data = sobj$data, gamma = 10^(-3:0), cost = 10^(-2:1)) 
# summary(tuned_parameters)
# figure best parameters and input into next
svm.model <- svm(as.factor(NOTCOV) ~ ., data = sobj$data, cost = 10, gamma = 0.1)
svm.pred <- predict(svm.model, s_dat_test)
table(pred = svm.pred, true = dat_test$NOTCOV)
```

# Model svm is showing that the model predicted People with health coverage are 60,212 wih error of 8,286, givig the percentage error equal: 12%
```{r}
8286/(60212+8286)
```
#While the predicted value for people who doesn't have coverage is 2,471 with error of 3,276, giving the percentage error equal: 57%
```{r}
3276/(3276+2471)
```

#Comparing the 4 models together Model_randfor has the lowest percentage error in predicting people who has health coverage  